---
title: "stringsAsFactors = HELLNO"
author: "Jenny Bryan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stringsAsFactors = HELLNO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Factors are a very useful type of variable in R, but they can also drive you nuts. Especially the "stealth factor" that you think of as character.

Can we soften some of their sharp edges?

Keeping order of levels via `givenorder()`:

```{r}
z <- factor(c("b", "c", "a", "e", "b"))

```

Normally the levels will be ordered alphabetically.

```{r}
factor(z)
```

Sometimes you'd like to maintain the order levels in which they appear in the data.

```{r}
givenorder(z)
```

Change some factors to characters

Check for factors where number of unique values = length, and change to character.

```{r}
m <- factor(c("a", "b", "c", "d"))
class(m)

m_character <- character(m)
class(m_character)
```

Otherwise print nochange.

```{r}
n <- factor(c("a", "b", "c", "d", "a"))
class(n)

character(n)

```

Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work.  Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```

